---
title: 'Prepare TOA matrix for YAPS'
output: 
  html_document:
    highlight: tango
    number_sections: false
    theme: united
    toc: yes
    toc_float:
      collapsed: false
  html_notebook:
    theme: united
    toc: yes
    toc_float:
      collapsed: false
    number_sections: false 
    highlight: tango
---
<style type="text/css">

body{ /* Normal  */
      font-size: 13px;
  }
td {  /* Table  */
  font-size: 11px;
}
h1.title {
  font-size: 26px;
}
h1 { /* Header 1 */
  font-size: 22px;
}
h2 { /* Header 2 */
    font-size: 16px;
}
h3 { /* Header 3 */
  font-size: 14px;
}
h4 { /* Header 3 */
  font-size: 9px;
}

</style>

```{r setup, include=FALSE}
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = as.numeric(difftime(Sys.time(), now, units = 'secs'))
      now <<- NULL
      # use options$label if you want the chunk label as well
      #paste('Chunk execution time: ', round(res, 2), ' sec')
      res0 <<- res
    }
  }})
)
```
<br>
The following methodology and function parameters are based directly on Jenna Vergeynst's [Python code](https://github.com/JennaVergeynst/prepare_toa_for_yaps).  
<mark style="background-color: #F2F2F2">Highlighted blocks</mark> are quotes from Jenna's repository.

Use this code to prepare [time-synchronized](https://elipickh.github.io/ReceiverArrays/notebooks/time_sync_jv_method.html) data from receiver arrays, which outputs a matrix, with one row per ping (with unrecorded pings set to NA) and one column per receiver.
The matrix can then be fed directly into [YAPS](https://github.com/baktoft/yaps) to get estimated locations. 

# Setup
* The <mark style="background-color: #F2F2F2"><span style="color:red">pkg_run</span></mark> function will load required libraries (and versions), installing/updating them first as needed
```{r}
# Disable scientific notation
options(scipen=999) 

# Print microseconds when viewing the data
sigFiguresSec = 20
options(digits.secs=sigFiguresSec)

# Import custom R functions
 for (f in list.files('../code', pattern='*.R')) {
     source(paste0('../code/', f))
 }

pkg_run(list(
            'lubridate', 
            c('data.table', '1.12.0'),
            'dygraphs',
            'htmltools'),
       load_verbose = FALSE,
       inst_verbose = TRUE)
```

# Load data
* Detection data from all receivers should be in one file
* Supply the timezone for the datetime column

The following sample data for [recs_df](https://github.com/JennaVergeynst/prepare_toa_for_yaps/blob/master/receiver_data_example.pkl) and [temperature](https://github.com/JennaVergeynst/prepare_toa_for_yaps/blob/master/temperature_data_example.pkl) were converted from pickle into csv.

```{r}
recs_df_path = '../data/recs_df_jv.csv'
temperature_path = '../data/temperature_jv.csv'

data_tz = 'GMT'
# same as data_tz = 'UTC'

Sys.setenv(TZ = data_tz)
```

* Read the df's and configure the datetime column
```{r}
recs_df = fread(recs_df_path, sep = ',', header = TRUE)[, 
          SyncTime := ymd_hms(SyncTime, tz = data_tz)]

temp_df = fread(temperature_path, sep = ',', header = TRUE)[, 
          Time := ymd_hms(Time, tz = data_tz)]
```

Total rows recs_df: ``r formatC(nrow(recs_df), , big.mark=",")``<br>
Total rows temp_df: ``r formatC(nrow(temp_df), , big.mark=",")``

```{r}
head(recs_df)
```

```{r}
head(temp_df)
```

* Specify the transmitter (i.e., fish) ID 
* Specify the datetime and receiver ID column names in `recs_df`
* Specify min and max burst for random burst interval transmitters  
```{r}
ID = 46903

dt_col = 'SyncTime'
rec_col = 'receiver_SN'

min_burst = 1.1
max_burst = 1.3
```

```{r}
setnames(temp_df, c(dt_col, 'temp_c'))
setkeyv(temp_df, dt_col)

mean_burst = round((min_burst + max_burst) / 2, 1)

# Subset recs_df to the given transmitter ID
tag_data = recs_df[Id == ID]
setkeyv(tag_data, dt_col)

rec_list = as.character(unique(tag_data$receiver_SN))
```

# Prepare TOA
* Join temperatue onto the detection data based on nearest matching time
* Calculate sound speed using Macknezie's formula (with salinity and depth fixed to 0 and 1, respectively)
```{r}
tag_data = temp_df[tag_data, roll = 'nearest'][, 
           soundspeed := soundspeed_seawater_mackenzie(temp = temp_c, salinity = 0, depth = 1)]
```

* Reshape detections data to one observation per row
* Split observations if time between detections are > <mark style="background-color: #F2F2F2"><span style="color:red">pas_tol</span></mark> minutes
* <mark style="background-color: #F2F2F2"><span style="color:red">max_time</span></mark>: <mark style="background-color: #F2F2F2">Maximum time (in seconds) that a sound signal would take between furthest receivers</mark>
```{r message=FALSE}
toa_data = prepare_tag_data(tag_data, dt_col, rec_col, rec_list, max_time = 0.2, pas_tol = 5)
```

# Clean TOA
* Clean up the TOA matrix:
  + <mark style="background-color: #F2F2F2">When observations follow eachother faster than the minimum known time delay, the observation with the least receivers is removed.</mark>
  + <mark style="background-color: #F2F2F2">Afterwards, the remaining observations that are still too close, are dropped if they have only one receiver in the row.</mark>
  + <mark style="background-color: #F2F2F2">Reasoning: these observations are too far apart to belong to the same row (i.e. further than max travel time between receivers), so are probably originating from a multipath.</mark>
```{r}
cleaned_toa_data = clean_toa_data(toa_data, 
                                  min_delay = (0.9 * min_burst), 
                                  rec_list)
```

* <mark style="background-color: #F2F2F2">Function to fill the gaps of missing pings in the dataframe, based on R code of Henrik Baktoft.</mark>
  + <mark style="background-color: #F2F2F2">Based on the average ping interval, a "virtual ping"-timeseries is created with stable burst interval on every average ping.</mark>
  + <mark style="background-color: #F2F2F2">Then the available pings are related to their closest virtual ping, and all virtual pings between get NaN-rows.</mark>
  + <mark style="background-color: #F2F2F2">only consider a track if it is more than 2 minutes long (120 seconds). Too short tracks will bug YAPS.</mark>
```{r}
filled_toa = lapply(split(cleaned_toa_data, by ='groups_pas'), 
                    function(x) fill_gaps(x, rec_list, dt_col, mean_burst, 
                                          min_track_length = 120))

filled_toa = filled_toa[!is.na(filled_toa)]
```

* Re-fill any new gaps which may have been created in the previous step
```{r}
final_toa = lapply(filled_toa,
                   function(x) fill_gaps(x, rec_list, dt_col, mean_burst, 
                                         min_track_length = 120))
```

# Plot TOA's
* Plot the resulting outputs
```{r}
ylim = c(min(toa_data$t_diff, na.rm = TRUE) - 1, 
         5*max_burst)
hlines = c(min_burst, max_burst)
height = 300
width = 800

p1 = dg_plot(toa_data, 
             main = paste0('Burst intervals fish ', ID, ' without cleaning'), 
             y_cols = 't_diff', dt_col = dt_col, line_w = 0, show_pts = TRUE, group = 'compare',
        ylim = ylim, hlines = hlines, height = height, width = width, r_selector = FALSE)

p2 = dg_plot(cleaned_toa_data, 
             main = paste0('Burst intervals fish ', ID, ' after 1st clean-up of too-close rows'), 
             y_cols = 't_diff', dt_col = dt_col, line_w = 0, show_pts = TRUE, group = 'compare',
        ylim = ylim, hlines = hlines, height = height, width = width, r_selector = FALSE)

p3 = dg_plot(dcast(rbindlist(filled_toa), get(dt_col) ~ groups_pas, value.var = 't_diff'), 
             main = paste0('Burst intervals fish ', ID, ' after filling of gaps'),
             group = 'compare', dt_col = 'get(dt_col)', line_w = 0, show_pts = TRUE,
        ylim = ylim, hlines = hlines, height = height, width = width, r_selector = FALSE)

p4 = dg_plot(dcast(rbindlist(final_toa), get(dt_col) ~ groups_pas, value.var = 't_diff'), 
             main = paste0('Burst intervals fish ', ID, ' after 2nd round of gap filling'), 
             group = 'compare', dt_col = 'get(dt_col)', line_w = 0, show_pts = TRUE,
        ylim = ylim, hlines = hlines, height = height, width = width, r_selector = FALSE)

browsable(htmltools::tagList(list(p1, p2, p3, p4)))
```

# Export final TOA
* Transform the outputs  

`toa_yaps` and `ss_data` can be fed into YAPS (along with recevier coordinates) to get estimated locations
```{r}
final_toa_df = setkeyv(rbindlist(final_toa), dt_col)

ss_data = final_toa_df$soundspeed
toa_yaps = t(as.matrix(final_toa_df[, rec_list, with=F]))
```

# Session Info

```{r}
sessionInfo()
```
